Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-11-05 writeups/Su_IGW_break.tex
stash work

The geometry of our simulation domain is fixed by one further parameter: $z_0$
the forcing location. We choose $L_z = 10H$ to give $\sim e^3$ amplitude growth
between the damping zones. We force at $z_0 = 2H = 0.2L_z$ within width
$\sigma = 0.078H$, sufficiently far from the lower damping zone and permit
sufficient room for the upward moving wave to grow as $\propto e^{z/2H}$.
Finally, we want similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$,
guided by the intuition that turbulence is approximately isotropic, so we use
$L_x = 4H$ and $N_z / N_x = 4$.

@@

The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
should be independent of $z$ between the forcing zone where it is generated and
the damping zone where it is dissipated (\autoref{eq:F_def}). We make may a
slightly more accurate estimate of the flux carried in the IGW by directly
substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this estimate
\begin{equation}
    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
\end{equation}
we may then measure the agreement of our simulation with analytical expectation
by computing
\begin{equation}
    \hat{F}(z, t) \equiv \frac{\ev{\rho u_xu_z}_x}
        {F_{al}'}.\label{eq:sbm_def}
\end{equation}
\autoref{fig:lin_fluxes} shows that $\hat{F}(z)\approx 1$ between $z_0$ and
$z_T$, as expected.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
    \caption{\autoref{eq:sbm_def} as a function of $z$ at select times $t$. As
    initial transients die out, $\hat{F}(z, t) = 1$ to very good agreement above
    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
    9.5H$. The flux excited in the forcing zone is transported without loss to
    the top of the domain, where it is dissipated by the damping layer (see
    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
\end{figure}

@@

In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $\hat{F}$ as a function
of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
$\overline{U}$ seems to conform qualitatively with the predictions of
\autoref{ss:crit_layer}, the behavior of $\hat{F}$ exhibits two notable
features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
there seems to be a small transmitted feature at many of the later times. We
will discuss further these features in \autoref{ss:reflectivity}, after first
analyzing the propagation of the critical layer.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
    the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
    1024$ simulation. $\overline{U}, \hat{F}$ follow their definitions in
    \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
of the critical layer. As the simulations are very noisy, we measure the location
of the critical layer using an average of where flux deposition occurs
\begin{align}
    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}'},\nonumber\\
    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}'},\nonumber\\
    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
\end{align}
This was found to a relatively stable estimator of the critical layer location.
Other estimators were used and do not significantly change the results of the
analysis.

@@

Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
0.71F_{al}'$ predictor, which serves mostly as a fiducial in the plot. The
general agreement clearly demonstrates $\Delta F < F_{al}'$ and that momentum
flux absorption is incomplete. However, its underprediction of critical layer
velocity at early times contrasts with its overprediction at later times. This
suggests that momentum flux absorption changes significantly over time, which is
in line with the observations in the subsequent \autoref{ss:reflectivity}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_front.png}
    \caption{Propagation of the critical layer over time. Shown are (black)
    $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
    integration of \autoref{eq:zc_anal} for $\Delta F(t)$ measured from
    simulation data (described in \autoref{eq:ds_def}) and (blue) direct
    substitution of time-averaged $\ev{\Delta F(t)}_t$ into
    \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
    conditions and integrate backwards, as $z_c$ is less well-defined at early
    times. The agreement of the directly-integrated predictor with the data
    shows \autoref{eq:zc_anal} is a good description of the evolution of $z_c$.
    The the poorer but qualitatively correct agreement of the time-averaged
    predictor with the data shows both that $\Delta F(t) \neq F_{al}'$ and that
    $\Delta F(t)$ likely has some very real variation with
    time.}\label{fig:nl_front}
\end{figure}

>>>>

The geometry of our simulation domain is fixed by one further parameter: $z_0$
the forcing location. We choose $L_z = 12.5H$ to give $\sim e^3$ amplitude growth
between the damping zones. We force at $z_0 = 3H$ within width $\sigma =
0.078H$, sufficiently far from the lower damping zone and permit sufficient room
for the upward moving wave to grow as $\propto e^{z/2H}$. Finally, we want
similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the
intuition that turbulence is approximately isotropic, so we use $L_x = 4H$ and
$N_z / N_x = 4$.

@@

The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
should be independent of $z$ between the forcing zone where it is generated and
the damping zone where it is dissipated (\autoref{eq:F_def}). We make may a
slightly more accurate estimate of the flux carried in the IGW by directly
substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this estimate
\begin{equation}
    F_{al}(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
\end{equation}
we may then measure the agreement of our simulation with analytical expectation
by computing
\begin{equation}
    F(z, t) = \frac{\ev{\rho u_xu_z}_x}.\label{eq:sbm_def}
\end{equation}
Where nonlinear effects are unimportant, we expect $F(z, t) = F_{al}(z)$.
\autoref{fig:lin_fluxes} shows that $F \approx F_{al}$ between $z_0$ and
$z_T$, as expected.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
    \caption{$F/F_{al}$ as a function of $z$ at select times $t$. As
    initial transients die out, $F / F_{al} = 1$ to very good agreement above
    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
    9.5H$. The flux excited in the forcing zone is transported without loss to
    the top of the domain, where it is dissipated by the damping layer (see
    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
\end{figure}

@@

In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
$\overline{U}$ seems to conform qualitatively with the predictions of
\autoref{ss:crit_layer}, the behavior of $F / F_{al}$ exhibits two notable
features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
there seems to be a small transmitted feature at many of the later times. We
will discuss further these features in \autoref{ss:reflectivity}, after first
analyzing the propagation of the critical layer.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $F(z, t) / F_{al}(z)$ respectively
    at the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
    1024$ simulation. $\overline{U}, F$ follow their definitions in
    \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
of the critical layer. As the simulations are very noisy, we measure the location
of the critical layer using an average of where flux deposition occurs
\begin{align}
    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
\end{align}
This was found to a relatively stable estimator of the critical layer location.
Other estimators were used and do not significantly change the results of the
analysis.

@@

Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
0.71F_{al}$ predictor, which serves mostly as a fiducial in the plot. The
general agreement clearly demonstrates $\Delta F < F_{al}$ and that momentum
flux absorption is incomplete. However, its underprediction of critical layer
velocity at early times contrasts with its overprediction at later times. This
suggests that momentum flux absorption changes significantly over time, which is
in line with the observations in the subsequent \autoref{ss:reflectivity}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_front.png}
    \caption{Propagation of the critical layer over time. Shown are (black)
    $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
    integration of \autoref{eq:zc_anal} for $\Delta F(t)$ measured from
    simulation data (described in \autoref{eq:ds_def}) and (blue) direct
    substitution of time-averaged $\ev{\Delta F(t)}_t$ into
    \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
    conditions and integrate backwards, as $z_c$ is less well-defined at early
    times. The agreement of the directly-integrated predictor with the data
    shows \autoref{eq:zc_anal} is a good description of the evolution of $z_c$.
    The the poorer but qualitatively correct agreement of the time-averaged
    predictor with the data shows both that $\Delta F(t) \neq F_{al}$ and that
    $\Delta F(t)$ likely has some very real variation with
    time.}\label{fig:nl_front}
\end{figure}
